---
title: "Coursera Statistical Inference Project Part 1"
author: "pankaj sharma"
date: "decemeber 24 2015"
output: pdf_document
---
#Synopsis

This is the project for the statistical inference coursera data science specialization class.This project consists of two parts :

* A simulation exercise.
* Basic inferential data analysis.

# Task

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. Set lambda = 0.2 for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. Note that you will need to do a thousand or so simulated averages of 40 exponentials.
```{r}
# set seed for reproducability
set.seed(31)

# set lambda to 0.2
lambda <- 0.2

# 40 samples
n <- 40

# 1000 simulations
simulations <- 1000

# simulate
simulated_exponentials <- replicate(simulations, rexp(n, lambda))

# calculate mean of exponentials
means_exponentials <- apply(simulated_exponentials, 2, mean)
```

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s. You should:

# Question 1
Show where the distribution is centered at and compare it to the theoretical center of the distribution.

```{r}
# distribution mean
analytical_mean <- mean(means_exponentials)
analytical_mean
```
```{r}
# analytical mean
theory_mean <- 1/lambda
theory_mean
```
```{r}
# visualization
hist(means_exponentials, xlab = "mean", main = "Exponential Function Simulations")
abline(v = analytical_mean, col = "red")
abline(v = theory_mean, col = "orange")

```


The analytics mean is 4.993867 the theoretical mean 5. The center of distribution of averages of 40 exponentials is very close to the theoretical center of the distribution.

# Question 2

Show how variable it is and compare it to the theoretical variance of the distribution.
```{r}
# standard deviation of distribution
standard_deviation_dist <- sd(means_exponentials)
standard_deviation_dist
```
```{r}
# standard deviation from analytical expression
standard_deviation_theory <- (1/lambda)/sqrt(n)
standard_deviation_theory
```
```{r}
# variance of distribution
variance_dist <- standard_deviation_dist^2
variance_dist
```
```{r}
# variance from analytical expression
variance_theory <- ((1/lambda)*(1/sqrt(n)))^2
variance_theory
```
Standard Deviation of the distribution is 0.7931608 with the theoretical SD calculated as 0.7905694. The Theoretical variance is calculated as 0.625. The actual variance of the distribution is 0.6291041


# Question3
Show that the distribution is approximately normal.

```{r}
xfit <- seq(min(means_exponentials), max(means_exponentials), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))
hist(means_exponentials,breaks=n,prob=T,col="green",xlab = "means",main="Density of means",ylab="density")
lines(xfit, yfit, pch=22, col="red", lty=5)

```
```{r}
# compare the distribution of averages of 40 exponentials to a normal distribution
qqnorm(means_exponentials,col=3)
qqline(means_exponentials, col = 2)
```


Due to Due to the central limit theorem (CLT), the distribution of averages of 40 exponentials is very close to a normal distribution.

