---
title: "Coursera Statistical Inference Project Part 2"
author: "pankaj sharma"
date: "decemeber 24 2015"
output: pdf_document
---

# 1. Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
library(datasets)
x <- ToothGrowth
# convert dose to factor
x$dose <- as.factor(x$dose)
```


```{r}
str(x)
```

```{r}
table(x$dose, x$supp)

```


```{r}
library(ggplot2)
ggplot(data=x, aes(x=dose, y=len, fill=supp)) + geom_bar(stat="identity",) + facet_grid(. ~ supp) + xlab("Dose in miligrams") + ylab("Tooth length") + guides(fill=guide_legend(title="Supplement type"))

```


```{r}
ggplot(aes(x=dose, y=len), data=x) + geom_boxplot(aes(fill=dose)) + xlab("Dose in miligrams") + ylab("Tooth length") + guides(fill=guide_legend(title="Dose"))

```

```{r}
ggplot(aes(x=supp, y=len), data=x) + geom_boxplot(aes(fill=supp)) + xlab("Supplement type") + ylab("Tooth length") + guides(fill=guide_legend(title="Supplement type"))

```

#  2. Provide a basic summary of the data.

```{r}
summary(x)

```
# 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
```{r}

t.test(len ~ supp, data = x)
```

Null hypothesis can not be rejected as confindence intervals contain zero and p-value is 0.06.

Supplement types seems to have no impact on Tooth growth

```{r}
# three groups as per dose level pairs
x.doses_0.5_1.0 <- subset (x, dose %in% c(0.5, 1.0)) 
x.doses_0.5_2.0 <- subset (x, dose %in% c(0.5, 2.0)) 
x.doses_1.0_2.0 <- subset (x, dose %in% c(1.0, 2.0)) 

# Check for dose levels (0.5, 1.0)
t.test(len ~ dose, data = x.doses_0.5_1.0)

```

```{r}
# Check for dose levels (0.5, 2.0)
t.test(len ~ dose, data = x.doses_0.5_2.0)

```

```{r}
# Check for dose levels (1.0, 2.0)
t.test(len ~ dose, data = x.doses_1.0_2.0)

```
The p-value is less than 0.05 and confindence intervals don't contian 0. The average toot length increases with an inceasing dose. The null hypothesis can be rejected.

# 4.1 Conclusions

* Supplement type seem to have no impact on tooth growth.
* Inreasing the dose level leads to increased tooth growth as well.

# 4.2 Assumptions

* The experiment was done with random assignment of guinea pigs to different dose level categories and supplement type to control for confounders that might affect the outcome.

* Members of the sample population, i.e. the 60 guinea pigs, are representative of the entire population of guinea pigs. This assumption allows us to generalize the results.

* For the t-tests, the variances are assumed to be different for the two groups being compared. This assumption is less stronger than the case in which the variances are assumed to be equal.


